Human-In-The-Loop Causal Discovery

### 2024-08-01 paper/main.tex
updates

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Title}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

Understanding cause-and-effect relationships between variables is a fundamental
goal in many scientific questions as it reveals the underlying mechanisms of
observed phenomena and informs effective interventions or policy decisions.
\emph{Causal Discovery} methods aim to uncover the underlying causal structure
among random variables from a given observational dataset. The problem of
causal discovery has been explored in both the Directed Acyclic Graphs (DAGs)
and Structural Equation Models (SEMs) literatures, each adopting different
approaches. In the DAG literature, the focus has primarily been on developing
automated algorithms that can learn the causal structure from observational
datasets. In contrast, the SEM literature has focused more towards developing
tools that assist researchers in manually constructing or modifying models.

@@

\begin{enumerate}
	\item \textbf{Make Obvious Mistakes:} Although most algorithms are
		theoretically consistent asymptotically, their finite sample
		properties are not well understood. Moreover, depending on the
		selected hyperparameters, their output can vary significantly.
		These issues combined with no good method to test the
		performance of these algorithms on a given dataset, in
		real-world settings with finite samples, these algorithms can
		make obvious mistakes. Consequently, researchers have to rely on
		posthoc modifications to the model based on their domain
		expertise.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
		DAGs can be faithful to a given observational dataset. Thus,
		using only observational data, algorithms can only recover the
		MECs that can contain a combination of directed and undirected
		edges. However, most methods for downstream tasks, such as
		identification or causal effect estimation, assume knowledge of
		a fully oriented DAG. As a result, researchers have to manually
		orient these edges before using them for analysis.
\end{enumerate}
\todo[inline]{The second point doesn't seem to be adding much}

@@

Inspired by this ranking approach in modification indices, in this paper we
develop a method to rank the CI tests using a conditional measure of
association. Our proposed approach uses the measure of association to find
correlated variables in the data that are not explained by the current model.
Using the measure, we can then rank the pair of variables between which adding
an edge would improve the explanation the most. Based on this information, the
user can select a potential pair of variables and based on their expert
knowledge decide the direction of the edge. Our main contributions in this
paper are as follows:
\begin{enumerate}
	\item We present a measure of (conditional) association for mixed data
		based on canonical correlations. This measure of association
		generalizes the widely used partial correlation to mixed data
		setting.
	\item Using this partial association measure, we present a procedure to
		rank CI tests. This ranking can be used to prioritize addition
		or deletion of edges from the model.
	\item Lastly, we present a web-browser based tool to allow researchers
		to apply this method to their own datasets.
\end{enumerate}

>>>>

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Human-In-The-Loop Causal Discovery}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

Understanding the cause-and-effect relationships between variables is a
fundamental goal in many scientific questions as it reveals the underlying
mechanisms of observed phenomena and informs effective interventions or policy
decisions. \emph{Causal Discovery} methods aim to uncover the underlying causal
structure among random variables from a given observational dataset. The
problem of causal discovery has been explored in both the Directed Acyclic
Graphs (DAGs) and Structural Equation Models (SEMs) literatures, each adopting
different approaches. In the DAG literature, the focus has primarily been on
developing automated algorithms that can learn the causal structure from
observational datasets. In contrast, the SEM literature has focused more
towards developing tools that assist researchers in manually constructing or
modifying models.

@@

\begin{enumerate}
	\item \textbf{Make Obvious Mistakes:} Although most algorithms are
		theoretically consistent asymptotically, their finite sample
		properties are not well understood. Moreover, depending on the
		selected hyperparameters, their output can vary significantly.
		These issues combined with no good method to test the
		performance of these algorithms on a given dataset, in
		real-world settings with finite samples, it can be difficult to
		trust the output of these algorithms. Consequently, researchers
		have to rely on posthoc modifications to the model based on
		their domain expertise.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
		DAGs can be faithful to a given observational dataset. Thus,
		using only observational data, algorithms can only recover the
		MECs that can contain a combination of directed and undirected
		edges. However, most methods for downstream tasks, such as
		identification or causal effect estimation, assume knowledge of
		a fully oriented DAG. As a result, researchers have to manually
		orient these edges before using them for analysis.
\end{enumerate}

\todo[inline]{Add an example here showing these problems}

@@

Inspired by this ranking approach in modification indices, in this paper we
develop a method to rank the CI tests using a conditional measure of
association. Our proposed approach uses the measure of association to find
correlated variables in the data that are not explained by the current model.
Using the measure, we can then rank the pair of variables between which adding
an edge would improve the explanation the most. Based on this information, the
user can select a potential pair of variables and based on their expert
knowledge decide the direction of the edge. 
\todo[inline]{Add about the score based methods somewhere here}

Our main contributions in this paper are as follows:
\begin{enumerate}
	\item We present a measure of (conditional) association for mixed data
		based on canonical correlations. This measure of association
		generalizes the widely used partial correlation to mixed data
		setting.
	\item Using this partial association measure, we present a procedure to
		rank CI tests. This ranking can be used to prioritize addition
		or deletion of edges from the model.
	\item Lastly, we present a web-browser based tool to allow researchers
		to apply this method to their own datasets.
\end{enumerate}
